### 2013-05-31 doc/dissertation/document/chapters/spn_equations/spn_equations.tex
updating spn relaxation section and data

For the ILUT, SPAINV, and SPAINV-ILUT preconditioning strategies the
reduced domain approximation will be applied to reduce the density of
the composite linear operator to more manageable levels. For each
preconditioner, the parameters that achieved the best quality metric
results from the previous analysis were used with ILUT parameters of a
fill level of 5.0 and a drop tolerance of \sn{1}{-5}, SPAINV
parameters of a sparsity level of 4 and a threshold of 0.1 and
SPAINV-ILUT with ILUT parameters of a fill level of 4.0 and a drop
tolerance of \sn{1}{-5} and SPAINV parameters of a sparsity level of 1
and a threshold of 1.0. The reduced domain threshold was set to
\sn{1}{-10} in order to eliminate any exceedingly small values from
the matrix (generally this is simply removing non-zero elements within
the floating point tolerance of zero). The reduced domain fill level
was then varied, starting with the largest non-zero entries per row
value observed for each of the preconditioning types in order to
assess its effects relative to the case where no reduced domain
approximation was applied.

Figure~\ref{fig:rda_iterations} gives the number of iterations
required to converge for each preconditioner type as a function of
reduced domain fill level. Figure~\ref{fig:rda_quality} gives the
corresponding quality metric for each data point where the number of
non-zero entries used to compute the metric is equivalent to the
reduced domain fill level. We first note the SPAINV preconditioning
alone is signficantly more sensitive to the reduction in domain size
over the ILUT-based methods, although the preconditioner was of
$O(100)$ non-zero entries per row without any approximation
applied. For the ILUT-based methods, performance was significantly
better with converge achieved in less than 40 MCSA iterations with
only 10 non-zero entries in each row (vs. 7 non-zero entries for the
case with no preconditioning). SPAINV-ILUT iterative performance was
marginally better than ILUT alone for all reduced domain fill
levels. However, it should be noted that the ILUT level of fill used
for the ILUT only calculations was set to 4 for this case while the
SPAINV-ILUT preconditioner used an ILUT level of fill of 5 and
therefore the marginally better performance is more likely a result of
this addition of fill rather than the extra SPAINV
preconditioning. Looking at the quality metric data, we see a nice
power-law reduction in the quality metric as a function of the reduced
domain fill level, achieving 2 orders of magnitude reduction in the
quality metric for the ILUT-based preconditioning methods.

Although applying the reduced domain approximation results in a
successful recovery of sparsity for the Monte Carlo problem while
maintaining good convergence properties, there is still an issue of
forming the composite operator before applying the approximation and
potentially generating the transpose of this operator in the case of
the adjoint Monte Carlo method. Because of this, memory and scaling
issues may still be observed when building the Monte Carlo problem. In
addition, the expensive extraction of the inverse of the
preconditioning operators for the explicit scheme creates a
significant cost in overall performance. Future work in this area
should be considerate of these important components of the algorithm.

@@

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/solver_iters.pdf}
  \end{center}
  \caption{\textbf{Average number of iterations required to converge
      each eigenvalue iteration for the fuel assembly problem as a
      function of energy groups.} \textit{All methods were
      preconditioned with identical ILUT paramters. For MCSA, several
      variations of fixed point iterations and estimators were used.}}
  \label{fig:comparison_iterations}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/solver_time.pdf}
  \end{center}
  \caption{\textbf{Average CPU time per iteration in seconds for the
      fuel assembly problem as a function of energy groups.}
    \textit{All linear solver iterations over all eigenvalue
      iterations were used to compute the averge. All methods were
      preconditioned with identical ILUT paramters. For MCSA, several
      variations of fixed point iterations and estimators were used.}}
  \label{fig:comparison_time}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/solver_p_time.pdf}
  \end{center}
  \caption{\textbf{Average CPU time per iteration in seconds for the
      fuel assembly problem as a function of energy groups with
      preconditioning time included for the MCSA methods.}
    \textit{All linear solver iterations over all eigenvalue
      iterations were used to compute the averge. All methods were
      preconditioned with identical ILUT paramters. For MCSA, several
      variations of fixed point iterations and estimators were used.}}
  \label{fig:comparison_prec_time}
\end{figure}

>>>>

For the ILUT, SPAINV, and SPAINV-ILUT preconditioning strategies the
reduced domain approximation will be applied to reduce the density of
the composite linear operator to more manageable levels. For each
preconditioner, the parameters that achieved the best quality metric
results from the previous analysis were used. These correspond to ILUT
parameters of a fill level of 5.0 and a drop tolerance of \sn{1}{-5},
SPAINV parameters of a sparsity level of 4 and a threshold of 0.1 and
SPAINV-ILUT with ILUT parameters of a fill level of 4.0 and a drop
tolerance of \sn{1}{-5} and SPAINV parameters of a sparsity level of 1
and a threshold of 1.0. The reduced domain threshold was set to
\sn{1}{-10} in order to eliminate any exceedingly small values from
the matrix (generally this is simply removing non-zero elements within
the floating point tolerance of zero). The reduced domain fill level
was then varied, starting with the largest non-zero entries per row
value observed for each of the preconditioning types in order to
assess its effects relative to the case where no reduced domain
approximation was applied.

Figure~\ref{fig:rda_iterations} gives the number of iterations
required to converge for each preconditioner type as a function of
reduced domain fill level. Figure~\ref{fig:rda_quality} gives the
corresponding quality metric for each data point where the number of
non-zero entries used to compute the metric is equivalent to the
reduced domain fill level. We first note the SPAINV preconditioning
alone is signficantly more sensitive to the reduction in domain size
over the ILUT-based methods, although the preconditioner was of
$O(100)$ non-zero entries per row without any approximation
applied. For the ILUT-based methods, performance was significantly
better with convergence achieved in less than 40 MCSA iterations with
only 10 non-zero entries in each row (vs. 7 non-zero entries for the
case with no preconditioning). SPAINV-ILUT iterative performance was
marginally better than ILUT alone for all reduced domain fill
levels. However, it should be noted that the ILUT level of fill used
for the ILUT only calculations was set to 4 for this case while the
SPAINV-ILUT preconditioner used an ILUT level of fill of 5 and
therefore the marginally better performance is more likely a result of
this addition of fill rather than the extra SPAINV
preconditioning. Looking at the quality metric data, we see a nice
power-law reduction in the quality metric as a function of the reduced
domain fill level, achieving 2 orders of magnitude reduction in the
quality metric for the ILUT-based preconditioning methods.

Although applying the reduced domain approximation results in a
successful recovery of sparsity for the Monte Carlo problem while
maintaining good convergence properties, there is still an issue of
forming the composite operator before applying the approximation and
potentially generating the transpose of this operator in the case of
the adjoint Monte Carlo method. Because of this, memory and scaling
issues may still be observed when building the probability and weight
matrices for the Monte Carlo problem. In addition, the expensive
extraction of the inverse of the preconditioning operators for the
explicit scheme creates a significant cost in overall
performance. Future work in this area should be considerate of these
important components of the preconditioned MCSA algorithm.

@@

We apply these relaxation parameters to the fuel assembly problem
along with the reduced domain approximation as a means of studying
their effects. For each calculation presented, the number of
iterations required to converge reported was for a single eigenvalue
iteration.  Again, \sn{3}{4} histories are used at each MCSA iteration
to compute the Monte Carlo correction using the adjoint collision
estimator. A reduced domain fill level of 100 is used with a threshold
of \sn{1}{-10} to filter small values.

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/rich_relax_iters.pdf}
  \end{center}
  \caption{\textbf{Number of iterations to converge a single
      eigenvalue iteration of the fuel assembly problem as a function
      of the Richardson relaxation parameter.} \textit{The Neumann
      relaxation parameter was fixed at 1.0 and \sn{3}{4} stochastic
      histories were used to compute the MCSA correction at each
      iteration.}}
  \label{fig:rich_relax_iters}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/rich_relax_time.pdf}
  \end{center}
  \caption{\textbf{CPU time in seconds to converge a single eigenvalue
      iteration of the fuel assembly problem as a function of the
      Richardson relaxation parameter.} \textit{The Neumann relaxation
      parameter was fixed at 1.0 and \sn{3}{4} stochastic histories
      were used to compute the MCSA correction at each iteration.}}
  \label{fig:rich_relax_time}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/neumann_relax_iters.pdf}
  \end{center}
  \caption{\textbf{Number of iterations to converge a single
      eigenvalue iteration of the fuel assembly problem as a function
      of the Neumann relaxation parameter.} \textit{The Richardson
      relaxation parameter was fixed at 1.0 and \sn{3}{4} stochastic
      histories were used to compute the MCSA correction at each
      iteration.}}
  \label{fig:neumann_relax_iters}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/neumann_relax_time.pdf}
  \end{center}
  \caption{\textbf{CPU time in seconds to converge a single eigenvalue
      iteration of the fuel assembly problem as a function of the
      Neumann relaxation parameter.} \textit{The Richardson relaxation
      parameter was fixed at 1.0 and \sn{3}{4} stochastic histories
      were used to compute the MCSA correction at each iteration.}}
  \label{fig:neumann_relax_time}
\end{figure}

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/solver_iters.pdf}
  \end{center}
  \caption{\textbf{Average number of iterations required to converge
      each eigenvalue iteration for the fuel assembly problem as a
      function of energy groups.} \textit{All methods were
      preconditioned with identical ILUT parameters. For MCSA, several
      variations of fixed point iterations and estimators were used.}}
  \label{fig:comparison_iterations}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/solver_time.pdf}
  \end{center}
  \caption{\textbf{Average CPU time per iteration in seconds for the
      fuel assembly problem as a function of energy groups.}
    \textit{All linear solver iterations over all eigenvalue
      iterations were used to compute the averge. All methods were
      preconditioned with identical ILUT parameters. For MCSA, several
      variations of fixed point iterations and estimators were used.}}
  \label{fig:comparison_time}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/solver_p_time.pdf}
  \end{center}
  \caption{\textbf{Average CPU time per iteration in seconds for the
      fuel assembly problem as a function of energy groups with
      preconditioning time included for the MCSA methods.}
    \textit{All linear solver iterations over all eigenvalue
      iterations were used to compute the averge. All methods were
      preconditioned with identical ILUT parameters. For MCSA, several
      variations of fixed point iterations and estimators were used.}}
  \label{fig:comparison_prec_time}
\end{figure}
